---
title: 'Getting and Cleaning Data: Course Project'
author: "Pietro Franzero"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---
  
```{r, echo = FALSE}
    library(knitr)
    library(rmarkdown)
    library(markdown)
```  
  
Before we begin, let's set echo to TRUE in the options, just to make sure the
code is easy to follow:
  
```{r, echo = TRUE}
    opts_chunk$set(echo = TRUE, results = "markup")
```
  
  
  
### Dowload and unzip file
  
First, we should download and unzip the data:
  
```{r}
    if(!file.exists("~/data")) {dir.create("~/data")}
    fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
    download.file(fileUrl, destfile = "~/data/database.zip")
    setwd("~/data")
    unzip("database.zip")
    setwd("~/data/UCI HAR Dataset")
```
  
  
  
### Load tables in parent folder
  
We can see that there are i) some text files, ii) a "train" folder, and iii) a "test" folder in the parent folder.
  
Let's load the features and activities tables and store them:
  
```{r}
    setwd("~/data/UCI HAR Dataset")
    
    features <- read.table("features.txt")

    activities <- read.table("activity_labels.txt")
    names(activities) <- c("activity_index","activity_label")
```
  
  
  
### Load tables from train folder
  
Now, let's load each table in the train folder, bind them and store results in a table "train":
  
```{r}
    setwd("~/data/UCI HAR Dataset/train")

    X_train <- read.table("X_train.txt")
    Y_train <- read.table("Y_train.txt")
    subject_train <- read.table("subject_train.txt")

    train <- cbind(subject_train,Y_train,X_train)
```
  
  
  
### Load tables from test folder
  
Let's do the same in the "test" folder and store results in a table "test":
  
```{r}
    setwd("~/data/UCI HAR Dataset/test")

    X_test <- read.table("X_test.txt")
    Y_test <- read.table("Y_test.txt")
    subject_test <- read.table("subject_test.txt")

    test <- cbind(subject_test,Y_test,X_test)
```
  
  
  
### Making the tidy dataset
  
After pre-processing the data, we are ready to make the tidy dataset:

**1.Merge the training and the test sets to create one data set.**
  
```{r}
    total <- rbind(train,test)
```
  
**4.Appropriately label the data set with descriptive variable names.**
  
```{r}
    features <- as.vector(features$V2)
    features <- append(features, c("Subject","Activity"), after = 0)
    names(total) <- features
```
  
**2.Extract only the measurements on the mean and standard deviation for each measurement.**
  
```{r}
    mean_and_std <- total[,which(grepl("mean()",names(total)) | grepl("std()",names(total)))]
    mean_and_std <- mean_and_std[,-which(grepl("-meanFreq()",names(mean_and_std)))]
    new_total <- cbind(Subject = total$Subject,Activity = total$Activity,mean_and_std)
```
  
  
**3. Use descriptive activity names to name the activities in the data set.**
  
```{r}
    for (i in 1:length(new_total$Activity)) {
      if (new_total$Activity[i] %in% activities$activity_index){
        new_total$Activity[i] <- as.character(activities[which(new_total$Activity[i] == activities$activity_index),]$activity_label)
      } else {
        stop("activity index not found")
      }
    }
```
  
**5. From the data set in step 4, create a second, independent tidy data set with the average of each variable for each activity and each subject.**
  
```{r}
    new_total$Subject <- as.factor(new_total$Subject)
    new_total$Activity <- as.factor(new_total$Activity)

    tidy <- aggregate(new_total[,-(1:2)],list(Subject = new_total$Subject, Activity = new_total$Activity), mean, na.rm = TRUE)
```
  
  
  
### Write a text file with tidy dataset
  
Now that we are done tidying the tidy dataset, we can store it in a text file:
  
```{r}
    setwd("~/data")
    write.table(tidy, file="tidy.txt", row.names=FALSE)
```
  